19Q2

### 2019-05-18 19Q2.Rmd
update Rmd

```{r merge_crispr_data, eval=TRUE}
crispr_all <- read.delim("./data_munging/19Q2/Achilles_gene_effect_20190514.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "DepMap_ID"

@@

```{r ccl_venn, eval=TRUE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr_ccls$DepMap_ID)

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists_data, eval=TRUE}
ccle_hists <- unique(select(ccle, DepMap_ID, disease)) %>% count(disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp, DepMap_ID, disease)) %>% count(disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc, DepMap_ID, disease)) %>% count(disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr, DepMap_ID, disease)) %>% count(disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists, fig.height=4.5, fig.width=4, eval=TRUE}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
png(filename = "./plots/manuscript/all_histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
dev.off()
```

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r, eval=TRUE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr$DepMap_ID)

@@

```{r wilcoxon_tests, eval=TRUE}
# crispr_ptmuts <- unfactorize(crispr_ptmuts)
# # crispr_signif_all_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_gene.rds")
# crispr_ptmuts_filt <- filter(crispr_ptmuts, Gene %in% g2p_genes)
# crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
# crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts_filt, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

# ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
# ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
# ccle_filt <- unfactorize(ccle_filt)
# ccle_signif_dg <- WilcoxonTestDF(dataframe = ccle, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_dg.rds")
# ccle_signif_g2p_dg <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_dg.rds")
# ccle_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ccle_signif_g2p_dgl.rds")

ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
ctrp_filt <- unfactorize(ctrp_filt)
ctrp_signif_dg <- WilcoxonTestDF(dataframe = ctrp, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_dg.rds")
ctrp_signif_g2p_dg <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_dg.rds")
ctrp_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ctrp_signif_g2p_dgl.rds")

gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
gdsc_filt <- unfactorize(gdsc_filt)
gdsc_signif_dg <- WilcoxonTestDF(dataframe = gdsc, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_dg.rds")
gdsc_signif_g2p_dg <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_dg.rds")
gdsc_signif_g2p_dgl <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/gdsc_signif_g2p_dgl.rds")
```

@@

```{r drug_spear_calc, warning=FALSE, eval=TRUE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

# CCLE
ccle_cor_res <- ccle %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_cor_res <- ctrp %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC
gdsc_cor_res <- gdsc %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

@@

Spearman correlations for gene expression and copy number:
```{r crispr_spear_calc, warning=FALSE}
crispr_spear_ge <- transform(crispr_ptmuts_filt %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr_ptmuts_filt %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
```

>>>>

```{r merge_crispr_data, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q2/Achilles_gene_effect_20190514.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "DepMap_ID"

@@

```{r ccl_venn, eval=FALSE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr$DepMap_ID)

@@

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r histology_dists_data, eval=FALSE}
ccle_hists <- unique(select(ccle, DepMap_ID, disease)) %>% count(disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp, DepMap_ID, disease)) %>% count(disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc, DepMap_ID, disease)) %>% count(disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr, DepMap_ID, disease)) %>% count(disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

```{r histology_dists, fig.height=4.5, fig.width=4, eval=FALSE}
drug_hists <- readRDS("./data_munging/rds/histology_distributions.rds")
png(filename = "./plots/manuscript/all_histology_distributions.png", width = 8, height = 9, units = "in", res = 500)
ggplot(drug_hists, aes(x = Dataset, y = n, fill = disease)) +
  geom_bar(stat = "identity", color = "black") +
  scale_y_continuous(breaks = seq(0, 850, by = 50), labels = seq(0, 850, by = 50)) +
  guides(fill = guide_legend(title = "Histology", ncol = 1)) +
  labs(y = "Count")
dev.off()
```

**RUN THIS WHEN CRISPR IS DONE MERGING (5/17)**
```{r, eval=FALSE}
ccle_ccls <- unique(ccle$DepMap_ID)
ctrp_ccls <- unique(ctrp$DepMap_ID)
gdsc_ccls <- unique(gdsc$DepMap_ID)
crispr_ccls <- unique(crispr$DepMap_ID)

@@

```{r wilcoxon_tests, eval=TRUE}
crispr <- unfactorize(crispr)
crispr_signif_all_gene <- WilcoxonTestDF(dataframe = crispr, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g.rds")
crispr_filt <- filter(crispr, Gene %in% g2p_genes)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_filt, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_g.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_filt, grouping = c("Gene", "disease"), outfile = "./data_munging/rds/crispr_signif_g2p_gl.rds")

# ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_genes)
# ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
# ccle_filt <- unfactorize(ccle_filt)
# ccle_signif_dg <- WilcoxonTestDF(dataframe = ccle, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_dg.rds")
# ccle_signif_g2p_dg <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_dg.rds")
# ccle_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ccle_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ccle_signif_g2p_dgl.rds")
# 
# ctrp_filt <- filter(ctrp, Drug %in% g2p_drugs & Gene %in% g2p_genes)
# ctrp_filt$Gene <- factor(ctrp_filt$Gene, ordered = FALSE)
# ctrp_filt <- unfactorize(ctrp_filt)
# ctrp_signif_dg <- WilcoxonTestDF(dataframe = ctrp, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_dg.rds")
# ctrp_signif_g2p_dg <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_dg.rds")
# ctrp_signif_g2p_dgl <- WilcoxonTestDF(dataframe = ctrp_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ctrp_signif_g2p_dgl.rds")
# 
# gdsc_filt <- filter(gdsc, Drug %in% g2p_drugs & Gene %in% g2p_genes)
# gdsc_filt$Gene <- factor(gdsc_filt$Gene, ordered = FALSE)
# gdsc_filt <- unfactorize(gdsc_filt)
# gdsc_signif_dg <- WilcoxonTestDF(dataframe = gdsc, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_dg.rds")
# gdsc_signif_g2p_dg <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_dg.rds")
# gdsc_signif_g2p_dgl <- WilcoxonTestDF(dataframe = gdsc_filt, grouping = c("Drug", "Gene", "disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/gdsc_signif_g2p_dgl.rds")
```

@@

```{r drug_spear_calc, warning=FALSE, eval=FALSE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

# CCLE
ccle_ge_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ccle_cn_res <- ccle_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, merge(ccle_ge_res, ccle_cn_res, by = c("Drug", "Gene", "Drug_Gene"), all = TRUE), by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_ge_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
ctrp_cn_res <- ctrp_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, merge(ctrp_ge_res, ctrp_cn_res, by = c("Drug", "Gene", "Drug_Gene"), all = TRUE), by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC
gdsc_ge_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(TPM) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "na.or.complete")$p.value)
gdsc_cn_res <- gdsc_filt %>% group_by(Drug, Gene, Drug_Gene) %>% drop_na(CN) %>% summarize(
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "na.or.complete")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, merge(gdsc_ge_res, gdsc_cn_res, by = c("Drug", "Gene", "Drug_Gene"), all = TRUE), by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

@@

Spearman correlations for gene expression and copy number:
```{r crispr_spear_calc, warning=FALSE}
crispr_spear_ge <- transform(crispr %>% group_by(Gene) %>% drop_na(TPM) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr %>% group_by(Gene) %>% drop_na(CN) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
c("white", "deepskyblue4")

ggplot(data = filter(crispr_spear_cn, Gene %in% g2p_genes)) +
  geom_point(aes(x = Gene, y = corr, fill = p_log10), pch = 21) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "bottom") +
  scale_fill_gradientn(colours = c("white", "green4"),
                       breaks = c(0, 4, 8, 12, 16, 20),
                       labels = c(0, 4, 8, 12, 16, 20),
                       limits = c(0, 21)) +
  geom_text_repel(data = filter(crispr_spear_cn, Gene %in% g2p_genes & p < 0.05) , mapping = aes(x = Gene, y = corr, label = Gene)) +
  labs(title = "Copy Number",
       fill = expression(-log[10](p)),
       x = "Gene",
       y = expression("Spearman Correlation Coefficient"~(r[s])))
```
